### 2015-11-13 projects/project_4/article/base.tex

\documentclass{article}

\usepackage[flushmargin, ragged, side]{footmisc}

\usepackage[]{amsmath} 
\usepackage[]{xcolor} 
\usepackage[]{amsthm} 
\usepackage[]{amssymb} 
\usepackage[]{enumerate} 
\usepackage[]{graphicx} 
\usepackage[]{subcaption} 
\usepackage[]{hyperref} 
\usepackage[]{cleveref}

\newcommand{\ua}{\uparrow}
\newcommand{\da}{\downarrow}
\newcommand*{\github}[1]{\def\@github{\color{gray!70}#1}}
\newcommand{\githublogo}{\raisebox{-1pt}{\includegraphics[height=9pt]{github}}}

\title{%
  \textsc{The Ising model in two dimensions}\\
  \small{FYS3150}
  }
\author{Ivar Haugal\o kken Stangeby}
\begin{document}
  \maketitle
  \begin{figure}[h!]
    \centering
    \includegraphics[width=0.8\linewidth]{new_york_skyline.pdf}
    \caption*{A Metropolis skyline as a function of temperature for 100000 MC-cycles.}

    \label{fig:new_york_skyline}
  \end{figure}
  \clearpage

  \begin{abstract}
    In this project we look at the Ising model in two dimensions. There are a
    set of physical quantities related to magnetic systems that are interesting
    to examine. We implement the Metropolis algorithm and compute various
    expectation-values for these quantities. For a $2\times2$ Ising model we
    derive the analytical expressions for these values and compare them with
    the numerically computed ones. We also examine what happen when we reach
    the thermodynamical limit, as the lattice size of the model increases.
    \noindent
    All source code can be found on my GitHub page: \\

    \centering{\href{https://github.com/qTipTip/FYS3150}{\githublogo \, \color{gray!50}\textit{github.com/qTipTip}}}
  \end{abstract}
  \tableofcontents
  % Explain the physical system we want to model
  % Introduce the Ising Model
  \input{ising_model.tex} 
  % - 1D model
  % - 2D model
  %     - Square lattice Ising Model
  %       - interacting vs noninteracting
  %       - with our without external magnetic field
  %       - we wish to lower energy, explain parallel vs anti-parallel spin
  % Introduce the Metropolis Algorithm and why we want to use it
  \input{metropolis_algorithm.tex}
  % implementation
  \input{implementations.tex}
  % simulations
  \input{simulations.tex}
  % Calculations of 2x2 analytical quantities
  \clearpage
  \input{conclusion.tex}
  \appendix
  \input{appendix_A_2x2Ising.tex}
\end{document}